---
title: "Master_thesis_final version with everything"
author: "Sara Lopez Fernandez"
date: "2023-05-01"
output: html_document
---
```{r packages}
library("dplyr")#manage data for plotting
library("magrittr")#to delete the 0
library("limma")
library(tidyverse) # great for averaging data and getting standard error for plotting # great for averaging data and getting standard error for plotting
library(vegan)
library(ggplot2)
library(phyloseq)
library(pheatmap)#give order to the heatmap

library(factoextra)#for the ordenation
library(ca)

library(reshape2)#to the heatmap, function melt
library(pals)#to the colours
library(plyr)#graph relative abundance reads and OTUs, sumarise
```

#Outputs 
##Rarefaction curves
```{r}
all_data <- read.table("C:/Users/salz0002/Documents/R/Data with outliers/CommunityData3.csv", header = TRUE, sep = ";")#data for rarefaction
```
```{r}
#compare x and y
all_data<-all_data %>% na.omit()
all_otu <- all_data[11:length(all_data)]
   
S_all <- specnumber(all_otu)  #observed number of species per sample
raremax <- min(rowSums(all_otu))
Srare <- rarefy(all_otu, raremax)
```
```{r}
plot(S_all, Srare, main="Sample", xlab = "Observed No. of Species", ylab = "Rarefied No. of Species")
```
```{r}
rarecurve(aggregate(all_otu, list(all_data$Site), sum)[,-1], xlab = "No. of reads", ylab = "No. of fungal OTUs", col = c("darkgreen", "orange", "red", "blue"), cex = 0.6) #without curve label

rarecurve(sapply(all_otu, function(x) sapply(split(x, all_data$Site), sum)), xlab = "No. of sequences", ylab = "No. of fungal OTUs") #with curve label
```

##Venn diagram plots

```{r}
data <- read.table("C:/Users/salz0002/Documents/R/Transdata.csv", header = TRUE, sep = ",") #data without outliers
data<-data %>% na.omit()
```
###Sites and treatment
```{r Split data based on site}
##Split the data or sorf data based on site
G_site <- data[data$Site=="G",]
g_otu<- subset(G_site, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs, PairY, PairT))
G_site <- colSums(g_otu)

M1_site <- data[data$Site=="M1",]
M1_otu<- subset(M1_site, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs,PairY, PairT))
M1_site <- colSums(M1_otu)

M2_site <- data[data$Site=="M2",]
M2_otu<- subset(M2_site, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs, PairY, PairT))
M2_site <- colSums(M2_otu)

#combine the three subdataframes
df <- cbind(G_site,M1_site,M2_site)
df <- as.data.frame(df)
df<-df[rowSums(df) !=0,]

#make binary the dataframe
G<-replace(df$G_site, df$G_site>0,1)
M1<-replace(df$M1_site, df$M1_site>0,1)
M2<-replace(df$M2_site, df$M2_site>0,1)

gsum <-  sum(G) #just to have a look at the number
m1sum <- sum(M1)
m2sum <- sum(M2)
 
df2 <- as.data.frame(cbind(G, M1, M2)) #only for 5 sets

#Draw the diagram
a <- vennCounts(df2)

vennDiagram(a)

##Treatment, natural or planted
#Split the data or sorf data based on site
N_treat <- data[data$Treatment=="N",]
N_otu<- subset(N_treat, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs, PairY, PairT))
N_treat <- colSums(N_otu)

P_treat <- data[data$Treatment=="P",]
P_otu<- subset(P_treat, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs, PairY, PairT))
P_treat <- colSums(P_otu)

#combine the three subdataframes
df <- cbind(N_treat,P_treat)
df <- as.data.frame(df)
df<-df[rowSums(df) !=0,]

#make binary the dataframe
N<-replace(df$N_treat, df$N_treat>0,1)
P<-replace(df$P_treat, df$P_treat>0,1)
Nsum <-  sum(N) #just to have a look at the number
Psum <- sum(P)
df2 <- as.data.frame(cbind(N, P)) #only for 5 sets

#Draw the diagram
b <- vennCounts(df2)
vennDiagram(b)
```
###Needle year
```{r}
##Split the data or sorf data based on site
Y21 <- data[data$Year=="21",]
Y21_otu<- subset(Y21, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs))
Y21 <- colSums(Y21_otu)

Y22<- data[data$Year=="22",]
Y22_otu<- subset(Y22, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs, PairY, PairT))
Y22 <- colSums(Y22_otu)

#combine the three subdataframes
library("dplyr")
df <- cbind(Y21,Y22)

df <- as.data.frame(df)

df<-df[rowSums(df) !=0,]

#make binary the dataframe

Y21<-replace(df$Y21, df$Y21>0,1)

Y22<-replace(df$Y22, df$Y22>0,1)

y21sum <-  sum(Y21) #just to have a look at the number

y22sum <- sum(Y22)

df2 <- as.data.frame(cbind(Y21, Y22)) #only for 5 sets

#Draw the diagram
library(limma)

c <- vennCounts(df2)

vennDiagram(c)
```


# Distribution over time
```{r}
datat <- read.table("C:/Users/salz0002/Documents/R/Data with outliers/CommunityData4(less).csv", header = TRUE, sep = ",")

datat<-datat %>% na.omit()
datat$Treatment <- as.factor(datat$Treatment)
datat$Year <- as.factor(datat$Year)
datat$week <- as.factor(datat$week)
```
```{r}
#G and natural
G_site <- datat[datat$Site=="G",]##Split the data or sorf data based on site
g_otu<- subset(G_site, select=-c(Tag, Site, Sample, Info, Time_point))
g_table<-subset(g_otu, select = -c(6:length(g_otu)))

G_natural <- g_table[g_table$Treatment=="N",]#seleccionar natural

 G_natural2 <-ddply(G_natural[,3:5], .(week), summarize, mean_reads=mean(reads))
 G_natural3 <-ddply(G_natural[,3:5], .(week), summarize, mean_OTUs=mean(OTUs))
 G_natural_fancy <- merge(G_natural2, G_natural3)
 
 
silly_vector <- rep("N", 8)
 G_natural_fancy <- cbind(silly_vector, G_natural_fancy)
```
```{r}
totalGN_reads<- sum(G_natural_fancy$mean_reads)#relative abundance for reads
rel_abundance<- G_natural_fancy$mean_reads/totalGN_reads
abundance_readsGN<- rel_abundance*100

totalGN_otus<-sum(G_natural_fancy$mean_OTUs)
otu_abun_GN<- G_natural_fancy$mean_OTUs/totalGN_otus
otu_GN_per<-otu_abun_GN*100

G_natural_real<-cbind(G_natural_fancy, abundance_readsGN, otu_GN_per)
```

```{r}
ggp<-ggplot(G_natural_real)+
  geom_bar(aes(x=week, y=abundance_readsGN), stat = "identity", fill="cyan",colour="#006000")+
  geom_point(aes(x=week, y=otu_GN_per),color="red", size=2)+
  geom_line(aes(x=week, y=otu_GN_per, group=1), stat="identity",color="blue", size=2 )+
  labs(title = "Relative abundance of natural regenerated seedlings in Gåvstahagen", x="weeks", y ="Relative abundance(%)")
ggp
```
```{r}
#G and Planted 
G_plant <- g_table[g_table$Treatment=="P",]
 G_plant2 <-ddply(G_plant[,3:5], .(week), summarize, mean_reads=mean(reads))
 G_plant3 <-ddply(G_plant[,3:5], .(week), summarize, mean_OTUs=mean(OTUs))
 G_plant_fancy <- merge(G_plant2, G_plant3)

silly_vector <- rep("P", 8)
 G_plant_fancy <- cbind(silly_vector, G_plant_fancy)
```
```{r}
totalGN_reads<- sum(G_plant_fancy$mean_reads)#relative abundance for reads
rel_abundance<- G_plant_fancy$mean_reads/totalGN_reads
abundance_readsGN<- rel_abundance*100

totalGN_otus<-sum(G_plant_fancy$mean_OTUs)
otu_abun_GN<- G_plant_fancy$mean_OTUs/totalGN_otus
otu_GN_per<-otu_abun_GN*100

G_plant_real<-cbind(G_plant_fancy, abundance_readsGN, otu_GN_per)
```

```{r}
ggp<-ggplot(G_plant_real)+
  geom_bar(aes(x=week, y=abundance_readsGN), stat = "identity", fill="cyan",colour="#006000")+
  geom_point(aes(x=week, y=otu_GN_per),color="red", size=2)+
  geom_line(aes(x=week, y=otu_GN_per, group=1), stat="identity",color="blue", size=2 )+
  labs(title = "Relative abundance of planted regenerated seedlings in Gåvstahagen", x="weeks", y ="Relative abundance(%)")
ggp

```
```{r}
#M1 and N
M1_site <- datat[datat$Site=="M1",]##Split the data or sorf data based on site
m1_otu<- subset(M1_site, select=-c(Tag, Site, Sample, Info, Time_point))
m1_table<-subset(m1_otu, select = -c(6:length(m1_otu)))

M1_natural <- m1_table[m1_table$Treatment=="N",]#seleccionar natural

M1_natural2 <-ddply(M1_natural[,3:5], .(week), summarize, mean_reads=mean(reads))
M1_natural3 <-ddply(M1_natural[,3:5], .(week), summarize, mean_OTUs=mean(OTUs))
M1_natural_fancy <- merge(M1_natural2, M1_natural3)
 
silly_vector <- rep("N", 11)
M1_natural_fancy <- cbind(silly_vector, M1_natural_fancy)
```
```{r}
totalGN_reads<- sum(M1_natural_fancy$mean_reads)#relative abundance for reads
rel_abundance<- M1_natural_fancy$mean_reads/totalGN_reads
abundance_readsGN<- rel_abundance*100

totalGN_otus<-sum(M1_natural_fancy$mean_OTUs)
otu_abun_GN<- M1_natural_fancy$mean_OTUs/totalGN_otus
otu_GN_per<-otu_abun_GN*100

M1_natural_real<-cbind(M1_natural_fancy, abundance_readsGN, otu_GN_per)
```

```{r}
ggp<-ggplot(M1_natural_real)+
  geom_bar(aes(x=week, y=abundance_readsGN), stat = "identity", fill="cyan",colour="#006000")+
  geom_point(aes(x=week, y=otu_GN_per),color="red", size=2)+
  geom_line(aes(x=week, y=otu_GN_per, group=1), stat="identity",color="blue", size=2 )+
  labs(title = "Relative abundance of natural regenerated seedlings in Marielund 1", x="weeks", y ="Relative abundance(%)")
ggp
```
```{r}
#M1 and P
M1_site <- data[data$Site=="M1",]##Split the data or sorf data based on site
m1_otu<- subset(M1_site, select=-c(Tag, Site, Sample, Info, Time_point))
m1_table<-subset(m1_otu, select = -c(6:length(m1_otu)))

M1_plant <- m1_table[m1_table$Treatment=="P",]#seleccionar natural

M1_plant2 <-ddply(M1_plant[,3:5], .(week), summarize, mean_reads=mean(reads))
M1_plant3 <-ddply(M1_plant[,3:5], .(week), summarize, mean_OTUs=mean(OTUs))
M1_plant_fancy <- merge(M1_plant2, M1_plant3)

silly_vector <- rep("P", 8)
M1_plant_fancy <- cbind(silly_vector, M1_plant_fancy)
```
```{r}
totalGN_reads<- sum(M1_plant_fancy$mean_reads)#relative abundance for reads
rel_abundance<- M1_plant_fancy$mean_reads/totalGN_reads
abundance_readsGN<- rel_abundance*100

totalGN_otus<-sum(M1_plant_fancy$mean_OTUs)
otu_abun_GN<- M1_plant_fancy$mean_OTUs/totalGN_otus
otu_GN_per<-otu_abun_GN*100

M1_plant_real<-cbind(M1_plant_fancy, abundance_readsGN, otu_GN_per)
```

```{r}
ggp<-ggplot(M1_plant_real)+
  geom_bar(aes(x=week, y=abundance_readsGN), stat = "identity", fill="cyan",colour="#006000")+
  geom_point(aes(x=week, y=otu_GN_per),color="red", size=2)+
  geom_line(aes(x=week, y=otu_GN_per, group=1), stat="identity",color="blue", size=2 )+
  labs(title = "Relative abundance of planted regenerated seedlings in Marielund 1", x="weeks", y ="Relative abundance(%)")
ggp
```

```{r}
#M2 and P
M2_site <- datat[datat$Site=="M2",]##Split the data or sorf data based on site
m2_otu<- subset(M2_site, select=-c(Tag, Site, Sample, Info, Time_point))
m2_table<-subset(m2_otu, select = -c(6:length(m2_otu)))

M2_plant <- m2_table[m2_table$Treatment=="P",]#seleccionar natural

M2_plant2 <-ddply(M2_plant[,3:5], .(week), summarize, mean_reads=mean(reads))
 M2_plant3 <-ddply(M2_plant[,3:5], .(week), summarize, mean_OTUs=mean(OTUs))
 M2_plant_fancy <- merge(M2_plant2, M2_plant3)
 
silly_vector <- rep("P", 7)
M2_plant_fancy <- cbind(silly_vector, M2_plant_fancy)
```
```{r}
totalGN_reads<- sum(M2_plant_fancy$mean_reads)#relative abundance for reads
rel_abundance<- M2_plant_fancy$mean_reads/totalGN_reads
abundance_readsGN<- rel_abundance*100

totalGN_otus<-sum(M2_plant_fancy$mean_OTUs)
otu_abun_GN<- M2_plant_fancy$mean_OTUs/totalGN_otus
otu_GN_per<-otu_abun_GN*100

M2_plant_real<-cbind(M2_plant_fancy, abundance_readsGN, otu_GN_per)
```

```{r}
ggp<-ggplot(M2_plant_real)+
  geom_bar(aes(x=week, y=abundance_readsGN), stat = "identity", fill="cyan",colour="#006000")+
  geom_point(aes(x=week, y=otu_GN_per),color="red", size=2)+
  geom_line(aes(x=week, y=otu_GN_per, group=1), stat="identity",color="blue", size=2 )+
  labs(title = "Relative abundance of planted regenerated seedlings in Marielund 2", x="weeks", y ="Relative abundance(%)")
ggp
```
```{r}
#M2 and N
M2_site <- datat[datat$Site=="M2",]##Split the data or sorf data based on site
m2_otu<- subset(M2_site, select=-c(Tag, Site, Sample, Info, Time_point))
m2_table<-subset(m2_otu, select = -c(6:length(m2_otu)))

M2_natural <- m2_table[m2_table$Treatment=="N",]#seleccionar natural

M2_natural2 <-ddply(M2_natural[,3:5], .(week), summarize, mean_reads=mean(reads))
 M2_natural3 <-ddply(M2_natural[,3:5], .(week), summarize, mean_OTUs=mean(OTUs))
 M2_natural_fancy <- merge(M2_natural2, M2_natural3)
 
silly_vector <- rep("P", 9)
M2_natural_fancy <- cbind(silly_vector, M2_natural_fancy)
```
```{r}
totalGN_reads<- sum(M2_natural_fancy$mean_reads)#relative abundance for reads
rel_abundance<- M2_natural_fancy$mean_reads/totalGN_reads
abundance_readsGN<- rel_abundance*100

totalGN_otus<-sum(M2_natural_fancy$mean_OTUs)
otu_abun_GN<- M2_natural_fancy$mean_OTUs/totalGN_otus
otu_GN_per<-otu_abun_GN*100

M2_natural_real<-cbind(M2_natural_fancy, abundance_readsGN, otu_GN_per)
```

```{r}
ggp<-ggplot(M2_natural_real)+
  geom_bar(aes(x=week, y=abundance_readsGN), stat = "identity", fill="cyan",colour="#006000")+
  geom_point(aes(x=week, y=otu_GN_per),color="red", size=2)+
  geom_line(aes(x=week, y=otu_GN_per, group=1), stat="identity",color="blue", size=2 )+
  labs(title = "Relative abundance of natural regenerated seedlings in Marielund 2", x="weeks", y ="Relative abundance(%)")
ggp
```


#CA
```{r}
data <- read.table("C:/Users/salz0002/Documents/R/Data with outliers/CommunityData4(less).csv", header = TRUE, sep = ",")

library("magrittr")

data<-data %>% na.omit()
```

```{r}
data$Site <- as.factor(data$Site)

data$Treatment <- as.factor(data$Treatment)

data$Year <- as.factor(data$Year)

data$week <- as.factor(data$week)

#str(all_data)
```

```{r deleting columns with 0}
n_otu<- data[11:length(data)]
n_otu2 <- n_otu[, colSums(n_otu != 0) > 0]

final_outliers<-cbind(data[,1:10],n_otu2)
```
```{r selecting the variable columns}
final_outliers<-final_outliers[,-1]
final_outliers<-final_outliers[,-5]
final_outliers<-final_outliers[,-6]
final_outliers<-final_outliers[,-6]
final_outliers<-final_outliers[,-6]
```
```{r getting coordinates for the ca}
final.ca<-ca(n_otu2, scale.=TRUE)
ca_dim1 <- fviz_contrib(final.ca, choice="col", axes=2, top = 20)
ca_dim1[["plot_env"]][["data_sum"]][["name"]]
ca_dim1
```
```{r}
###Distintas formas de hacer el CA
ca1<-fviz_ca_row(final.ca, col.row = "cos2", gradient.col = c("#00AFBB", "#E7B800", "#FC4E07"))
ca2<- fviz_ca_col(final.ca, col.row = "cos2", gradient.col = c("#00AFBB", "#E7B800", "#FC4E07"))
ca3<-fviz_ca_col(final.ca, col.row = "cos2", gradient.col = c("#00AFBB", "#E7B800", "#FC4E07"))
ca4<-fviz_ca_biplot(final.ca, col.row = "cos2", gradient.col = c("#00AFBB", "#E7B800", "#FC4E07"))
ca5<-fviz_ca(final.ca, col.row = "cos2", gradient.col = c("#00AFBB", "#E7B800", "#FC4E07"))

```
####CA with outliers and labels
```{r the Ca1 with the outliers}
ca1_coord.df <- ca1$data
View(ca1_coord.df)

meta<-final_outliers[, 1:5]
ca1_df<- cbind(meta,ca1_coord.df)

library(ggplot2)
library(RColorBrewer)

#Sample
ca1_sam <- ggplot(ca1_df, aes(x=x, y=y, colour = cos2, label = Sample)) + geom_point() + geom_text(hjust = 0, vjust = 0) + labs(x = "Dim1 (6.3%)", y = "Dim2(5.7%)")
```

```{r the Ca2 with the outliers otus}
ca2_coord.df <- ca2$data

#Sample
ca2_sam <- ggplot(ca2_coord.df, aes(x=x, y=y, colour = cos2, label = name)) + geom_point() + geom_text(hjust = 0, vjust = 0)+ labs(x = "Dim1 (6.3%)", y = "Dim2(5.7%)")
```

```{r without outliers}
new_data<-data[-c(48,70),]
```
```{r}
n_otu<- new_data[11:length(new_data)]
n_otu2 <- n_otu[, colSums(n_otu != 0) > 0]

final<-cbind(new_data[,1:10],n_otu2)
```
```{r selecting the variable columns}
final<-final[,-1]
final<-final[,-5]
final<-final[,-6]
final<-final[,-6]
final<-final[,-6]
```
```{r getting coordinates for the ca}
final.ca<-ca(n_otu2, scale.=TRUE)
ca_dim1 <- fviz_contrib(final.ca, choice="col", axes=2, top = 20)
ca_dim1[["plot_env"]][["data_sum"]][["name"]]
ca_dim1
```
```{r}
ca6<-fviz_ca_row(final.ca, col.row = "cos2", gradient.col = c("#00AFBB", "#E7B800", "#FC4E07"))
ca7<- fviz_ca_col(final.ca, col.row = "cos2", gradient.col = c("#00AFBB", "#E7B800", "#FC4E07"))

```


###CA without outliers
```{r without outliers}
new_data<-data[-c(48,70),]
```
```{r}
n_otu<- new_data[11:length(new_data)]
n_otu2 <- n_otu[, colSums(n_otu != 0) > 0]

final<-cbind(new_data[,1:10],n_otu2)
```
```{r selecting the variable columns}
final<-final[,-1]
final<-final[,-5]
final<-final[,-6]
final<-final[,-6]
final<-final[,-6]
```
```{r getting coordinates for the ca}
final.ca<-ca(n_otu2, scale.=TRUE)
ca_dim1 <- fviz_contrib(final.ca, choice="col", axes=2, top = 20)
ca_dim1[["plot_env"]][["data_sum"]][["name"]]
ca_dim1
```
```{r}
ca6<-fviz_ca_row(final.ca, col.row = "cos2", gradient.col = c("#00AFBB", "#E7B800", "#FC4E07"))
ca7<- fviz_ca_col(final.ca, col.row = "cos2", gradient.col = c("#00AFBB", "#E7B800", "#FC4E07"))

```
####Final CA without outliers and labels
```{r select the coordinates to perform the ggplot}
ca6_coord.df <- ca6$data
View(ca6_coord.df)

meta<-final[, 1:5]
ca6_df<- cbind(meta,ca6_coord.df)
```
```{r}
library(ggplot2)
library(RColorBrewer)

#week
colors <- rainbow(length(ca6_df$week))
ca6_week <- ggplot(ca6_df, aes(x=x, y=y, colour = week, label = Sample)) + geom_point() + scale_fill_manual(values = colors) + geom_text(hjust = 0, vjust = 0) + labs(x = "Dim1 (6.3%)", y = "Dim2(5.7%)")

#year
colors <- rainbow(length(ca6_df$Year))
ca6_week <- ggplot(ca6_df, aes(x=x, y=y, colour = Year, label = Sample)) + geom_point() + scale_fill_manual(values = colors) + geom_text(hjust = 0, vjust = 0) + labs(x = "Dim1 (6.3%)", y = "Dim2(5.7%)")

#treatment
colors <- rainbow(length(ca6_df$Treatment))
ca6_week <- ggplot(ca6_df, aes(x=x, y=y, colour = Treatment, label = Sample)) + geom_point() + scale_fill_manual(values = colors) + geom_text(hjust = 0, vjust = 0) + labs(x = "Dim1 (6.3%)", y = "Dim2(5.7%)")

#site
colors <- rainbow(length(ca6_df$Site))
ca6_week <- ggplot(ca6_df, aes(x=x, y=y, colour = Site, label = Sample)) + geom_point() + scale_fill_manual(values = colors) + geom_text(hjust = 0, vjust = 0) + labs(x = "Dim1 (6.3%)", y = "Dim2(5.7%)")
```
```{r the Ca2 with the outliers otus}
ca7_coord.df <- ca7$data

#Sample
ca7_sam <- ggplot(ca7_coord.df, aes(x=x, y=y, colour = cos2, label = name)) + geom_point() + geom_text(hjust = 0, vjust = 0)+ labs(x = "Dim1 (6.3%)", y = "Dim2(5.7%)")
```


#Histograms-->Normality
```{r}
h1<-hist(data$reads)

h2<-hist(data$OTUs)
```
##Contraste de hipótesis: normalidad
```{r}
shapiro.test(data$OTUs)
shapiro.test(data$reads)
```
##homocedasticidad
```{r}
fligner.test(OTUs~Treatment, data)

fligner.test(OTUs~Site, data)

fligner.test(OTUs~Year, data)

fligner.test(OTUs~week, data)


fligner.test(reads~Treatment, data)

fligner.test(reads~Site, data)

fligner.test(reads~Year, data)

fligner.test(reads~week, data)###Mis datos son normales y homocedasticos
##Heterodascity. The variability is equally distributed--> it doesn't matter the value of the residual or the real value.
```


#Diversity analysis
```{r}
data <- read.table("C:/Users/salz0002/Documents/R/Transdata.csv", header = TRUE, sep = ",")

n_otu <- data[11:length(data)]
n_meta <- subset(data[1:10])
str(n_meta)

n_meta$Treatment <- as.factor(n_meta$Treatment)
n_meta$Site <- as.factor(n_meta$Site)
n_meta$week <- as.factor(n_meta$week)
n_meta$Year <- as.factor(n_meta$Year)

all_data <- cbind(n_meta, n_otu)


###Normalised data with read count size.Root square of each OTU per sample

all_data<-cbind(all_data, sqrt(rowSums(n_otu)))
colnames(all_data)[colnames(all_data) =="sqrt(rowSums(n_otu))"]<- "read.counts"
```
##Alfa-diversity
```{r model with all the variables}
#alfa-diversty --> Shannon
shann <- diversity(n_otu, "shannon")

all_data$Shannon<-shann #add a column with the shannon data

###models
#all the variables
H<-lm(shann~read.counts+Treatment+Site+Year+week, data = all_data)
H
one<- aov(H, data = all_data)

summary(one)
plot(one)
```
```{r post hoc alfa-diversity}
##pairwise comparation post hoc alfa-diversity para weeks
pairwise.t.test(shann, all_data$week, p.adj = "bonferroni")#bonferroni es exigente no salen cosas
TDH<-data.frame(TukeyHSD(one, which = "week", conf.level = 0.95))##este me cae mejor salen signif. 
```
###Boxplot alfa-diversity
```{r}
bp<-boxplot(Shannon~Treatment, all_data, xlab = "Seedling type",
   ylab = "Shannon index", main = "Alpha diversity based on seedling type")
```
```{r boxplot shannon-treatment}
St<-boxplot(Shannon ~ Treatment, data = all_data, 
   xlab = "Seedling type",
   ylab = "Shannon index", 
   main = "Alpha diversity based on the seedling type",
   notch = TRUE, 
   varwidth = TRUE, 
   col = c("#00BA42", "#ED8141"))#,names = c("N","P")
```
```{r boxplot shannon-site}
St<-boxplot(Shannon ~ Site, data = all_data, 
   xlab = "Site",
   ylab = "Shannon index", 
   main = "Alpha diversity based on the site",
   notch = TRUE,
   varwidth = TRUE, 
   col = c("#00BA42", "#ED8141", "#BF80FF"))#,names = c("N","P")
```
```{rboxplot shannon-year}
St<-boxplot(Shannon ~ Year, data = all_data, 
   xlab = "Year",
   ylab = "Shannon index", 
   main = "Alpha diversity based on the needle age",
    notch = TRUE,
   varwidth = TRUE, 
   col = c("#00BA42", "#BF80FF"),
   names = c("Y21","Y22"))
```
```{rboxplot shannon-week}
St<-graphics::boxplot(Shannon ~ week, data = all_data, 
   xlab = "Weeks",
   ylab = "Shannon index", 
   main = "Alpha diversity based on the exposure time",
   varwidth = TRUE, 
   col = c("lightgreen"))
```


##Beta diversity
```{r beta diversity}
####Beta diversity.
#I need to transform the data with Hellinger transformation
library("labdsv")
hell<-hellinger(n_otu)

#All variable
bd<-adonis2(hell~read.counts+Treatment+Site+Year+week, data = all_data, method = "bray")
bd
summary(bd)
```
###Heatmap plot
```{r Importing taxa and all data files}
data <- read.table("C:/Users/salz0002/Documents/R/Transdata.csv", header = TRUE, sep = ",")#less OTUs and no outliers
```
```{r data preparation}
#select only the OTUs
otutb<-data[13:824]
```

##Paired-wise comparation-->Hannah´s paper
```{r data}
data <- read.table("C:/Users/salz0002/Documents/R/Transdata.csv", header = TRUE, sep = ",")

library("magrittr")

data<-data %>% na.omit()
```
```{r preparing data as factors}
data$Site <- as.factor(data$Site)
data$Treatment <- as.factor(data$Treatment)

data$Year <- as.factor(data$Year)

data$week <- as.factor(data$week)

data$PairT<-as.factor(data$PairT)
data$PairY<- as.factor(data$PairY)
```
```{r adding the read counts in the table}
n_otu<-data[,13:length(data)]
data<-cbind(data, sqrt(rowSums(n_otu)))
colnames(data)[colnames(data) =="sqrt(rowSums(n_otu))"]<- "read.counts"
meta<-data[,c(3:5,7,9,10,834)]
```
*RICHNESS - Hannah's way but applying a lineal model*
```{r Shannon, alpha diversity}
tb<- cbind(meta,n_otu)

shanntb <- diversity(tb[,8:828], "shannon")
```

```{r model with the pair T}
test1<- aov(shanntb ~ read.counts + PairT, data = tb)
summary(test1)
```

```{r Model with the Pair Y}
test2<- aov(shanntb ~ read.counts + PairY, data = tb)
summary(test2)
```


#Relative abundance with the taxa (20 tops otus)
##Ra graph for the Select taxa over weeks
```{r}
data <- read.table("C:/Users/salz0002/Documents/R/Rabundance_20topsotu.csv", header = TRUE, sep = ",")
```
```{r}
data$week <- as.factor(data$week)
```
```{r}
dfm= reshape2::melt(data, na.rm = FALSE, value.name = "value")
```
```{r color object}
colours=c("royalblue","brown2", "lightgreen", "magenta","seagreen","chocolate1", "cyan", "darkgoldenrod2", "darkmagenta", "plum", "snow", "olivedrab", "lightsalmon","lightsalmon4", "lightblue","slateblue", "plum4", "wheat3", "blue", "yellow", "slategray4")
```
```{r plot}
tax20 = ggplot(dfm, aes(x = week, fill = variable, y = value)) + 
    geom_bar(stat = "identity", colour = "black") + 
    theme(axis.text.x = element_text(angle = 90, size = 14, colour = "black", vjust = 0.5, hjust = 1, face= "bold"), 
    axis.title.y = element_text(size = 16, face = "bold"),  axis.title.x = element_text(size = 16, face = "bold"), legend.title = element_text(size = 16, face = "bold"),     panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
    legend.text = element_text(size = 12, face = "bold", colour = "black"), 
    axis.text.y = element_text(colour = "black", size = 12, face = "bold")) + 
    scale_y_continuous(expand = c(0,0)) + 
    labs(x = "Weeks", y = "Relative Abundance (%)", fill = "20 most abundant OTUs") + 
    scale_fill_manual(values = colours)
    
```

##Ra graph PairY
###Gavsta
```{r}
G <- read.table("C:/Users/salz0002/Documents/R/Paired sample design/G_PairY_RA.csv", header = TRUE, sep = ",")

G<-G[, 2:length(G)]
```
```{r}
G$OTUs <- factor(G$OTUs, levels=G$OTUs)
```
```{r}
dfm= melt(G, na.rm = FALSE, value.name = "value")
```


```{r}
colours=c("royalblue","brown2", "lightgreen", "magenta","seagreen","chocolate1", "cyan", "darkgoldenrod2", "darkmagenta", "plum", "snow", "olivedrab", "lightsalmon","lightsalmon4", "lightblue","slateblue", "plum4", "wheat3", "blue", "yellow", "slategray4")
```

```{r}
tax20 = ggplot(dfm, aes(x = OTUs, fill = variable, y = value)) + 
    geom_bar(stat = "identity", colour = "black")+
    theme(axis.text.x = element_text(angle = 90, size = 14, colour = "black", vjust = 0.5, hjust = 1, face= "bold"), 
    axis.title.y = element_text(size = 16, face = "bold"),  axis.title.x = element_text(size = 16, face = "bold"), legend.title = element_text(size = 16, face = "bold"),     panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
    legend.text = element_text(size = 12, face = "bold", colour = "black"), 
    axis.text.y = element_text(colour = "black", size = 12, face = "bold")) + 
    scale_y_continuous(expand = c(0,0)) + 
    labs(x = "Pair year over time", y = "Relative Abundance (%)", fill = "20 most abundant OTUs") + 
    scale_fill_manual(values = colours)+
    ggtitle("Gåvstahagen- Paired samples based on needle age")
```
###M1
```{r}
M1 <- read.table("C:/Users/salz0002/Documents/R/Paired sample design/M1_PairY_RA.csv", header = TRUE, sep = ",")

M1<-M1[, 2:length(M1)]
```
```{r}
M1$OTUs <- factor(M1$OTUs, levels=M1$OTUs)
```
```{r}
dfm= melt(M1, na.rm = FALSE, value.name = "value")
```


```{r}
colours=c("royalblue","brown2", "lightgreen", "magenta","seagreen","chocolate1", "cyan", "darkgoldenrod2", "darkmagenta", "plum", "snow", "olivedrab", "lightsalmon","lightsalmon4", "lightblue","slateblue", "plum4", "wheat3", "blue", "yellow", "slategray4")
```

```{r}
tax20 = ggplot(dfm, aes(x = OTUs, fill = variable, y = value)) + 
    geom_bar(stat = "identity", colour = "black")+
    theme(axis.text.x = element_text(angle = 90, size = 14, colour = "black", vjust = 0.5, hjust = 1, face= "bold"), 
    axis.title.y = element_text(size = 16, face = "bold"),  axis.title.x = element_text(size = 16, face = "bold"), legend.title = element_text(size = 16, face = "bold"),     panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
    legend.text = element_text(size = 12, face = "bold", colour = "black"), 
    axis.text.y = element_text(colour = "black", size = 12, face = "bold")) + 
    scale_y_continuous(expand = c(0,0)) + 
    labs(x = "Pair year over time", y = "Relative Abundance (%)", fill = "20 most abundant OTUs") + 
    scale_fill_manual(values = colours)+
    ggtitle("Marielund 1- Paired samples based on needle age")
   
```
###M2
```{r}
M2 <- read.table("C:/Users/salz0002/Documents/R/Paired sample design/M2_PairY_RA.csv", header = TRUE, sep = ",")

M2<-M2[, 2:length(M2)]
```
```{r}
M2$OTUs <- factor(M2$OTUs, levels=M2$OTUs)
```
```{r}
dfm= melt(M2, na.rm = FALSE, value.name = "value")
```


```{r}
colours=c("royalblue","brown2", "lightgreen", "magenta","seagreen","chocolate1", "cyan", "darkgoldenrod2", "darkmagenta", "plum", "snow", "olivedrab", "lightsalmon","lightsalmon4", "lightblue","slateblue", "plum4", "wheat3", "blue", "yellow", "slategray4")
```

```{r}
tax20 = ggplot(dfm, aes(x = OTUs, fill = variable, y = value)) + 
    geom_bar(stat = "identity", colour = "black")+
    theme(axis.text.x = element_text(angle = 90, size = 14, colour = "black", vjust = 0.5, hjust = 1, face= "bold"), 
    axis.title.y = element_text(size = 16, face = "bold"),  axis.title.x = element_text(size = 16, face = "bold"), legend.title = element_text(size = 16, face = "bold"),     panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
    legend.text = element_text(size = 12, face = "bold", colour = "black"), 
    axis.text.y = element_text(colour = "black", size = 12, face = "bold")) + 
    scale_y_continuous(expand = c(0,0)) + 
    labs(x = "Pair year over time", y = "Relative Abundance (%)", fill = "20 most abundant OTUs") + 
    scale_fill_manual(values = colours)+
  ggtitle("Marielund 2-  Paired samples based on needle age")
   
```

##Ra graph PairTreatment 
##Gavsta 

```{r}
G <- read.table("C:/Users/salz0002/Documents/R/Paired sample design/G_PairT.Porcenta.csv", header = TRUE, sep = ",")
```
```{r}
G$OTUs <- factor(G$OTUs, levels=G$OTUs)
```
```{r}
dfm= melt(G, na.rm = FALSE, value.name = "value")
```


```{r}
colours=c("royalblue","brown2", "lightgreen", "magenta","seagreen","chocolate1", "cyan", "darkgoldenrod2", "darkmagenta", "plum", "snow", "olivedrab", "lightsalmon","lightsalmon4", "lightblue","slateblue", "plum4", "wheat3", "blue", "yellow", "slategray4")
```

```{r}
tax20 = ggplot(dfm, aes(x = OTUs, fill = variable, y = value)) + 
    geom_bar(stat = "identity", colour = "black")+
    theme(axis.text.x = element_text(angle = 90, size = 14, colour = "black", vjust = 0.5, hjust = 1, face= "bold"), 
    axis.title.y = element_text(size = 16, face = "bold"),  axis.title.x = element_text(size = 16, face = "bold"), legend.title = element_text(size = 16, face = "bold"),     panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
    legend.text = element_text(size = 12, face = "bold", colour = "black"), 
    axis.text.y = element_text(colour = "black", size = 12, face = "bold")) + 
    scale_y_continuous(expand = c(0,0)) + 
    labs(x = "Pair treatment over time", y = "Relative Abundance (%)", fill = "20 most abundance OTUs") + 
    scale_fill_manual(values = colours)+
  ggtitle("Gåvstahagen- Paired samples based on seedling type")
```
##M1
```{r}
M1 <- read.table("C:/Users/salz0002/Documents/R/Paired sample design/M1_PairT.Porcenta.csv", header = TRUE, sep = ",")
```
```{r}
M1$OTUs <- factor(M1$OTUs, levels=M1$OTUs)
```
```{r}
dfm= melt(M1, na.rm = FALSE, value.name = "value")
```


```{r}
colours=c("royalblue","brown2", "lightgreen", "magenta","seagreen","chocolate1", "cyan", "darkgoldenrod2", "darkmagenta", "plum", "snow", "olivedrab", "lightsalmon","lightsalmon4", "lightblue","slateblue", "plum4", "wheat3", "blue", "yellow", "slategray4")
```

```{r}
tax20 = ggplot(dfm, aes(x = OTUs, fill = variable, y = value)) + 
    geom_bar(stat = "identity", colour = "black")+
    theme(axis.text.x = element_text(angle = 90, size = 14, colour = "black", vjust = 0.5, hjust = 1, face= "bold"), 
    axis.title.y = element_text(size = 16, face = "bold"),  axis.title.x = element_text(size = 16, face = "bold"), legend.title = element_text(size = 16, face = "bold"),     panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
    legend.text = element_text(size = 12, face = "bold", colour = "black"), 
    axis.text.y = element_text(colour = "black", size = 12, face = "bold")) + 
    scale_y_continuous(expand = c(0,0)) + 
    labs(x = "Pair treatment over time", y = "Relative Abundance (%)", fill = "20 most abundance OTUs") + 
    scale_fill_manual(values = colours)+
  ggtitle("Marielund 1- Paired samples based on seedling type")
   
```
##M2
```{r}
M2 <- read.table("C:/Users/salz0002/Documents/R/Paired sample design/M2_PairT.Porcenta.csv", header = TRUE, sep = ",")
```
```{r}
M2$OTUs <- factor(M2$OTUs, levels=M2$OTUs)
```
```{r}
dfm= melt(M2, na.rm = FALSE, value.name = "value")
```


```{r}
colours=c("royalblue","brown2", "lightgreen", "magenta","seagreen","chocolate1", "cyan", "darkgoldenrod2", "darkmagenta", "plum", "snow", "olivedrab", "lightsalmon","lightsalmon4", "lightblue","slateblue", "plum4", "wheat3", "blue", "yellow", "slategray4")
```

```{r}
tax20 = ggplot(dfm, aes(x = OTUs, fill = variable, y = value)) + 
    geom_bar(stat = "identity", colour = "black")+
    theme(axis.text.x = element_text(angle = 90, size = 14, colour = "black", vjust = 0.5, hjust = 1, face= "bold"), 
    axis.title.y = element_text(size = 16, face = "bold"),  axis.title.x = element_text(size = 16, face = "bold"), legend.title = element_text(size = 16, face = "bold"),     panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
    legend.text = element_text(size = 12, face = "bold", colour = "black"), 
    axis.text.y = element_text(colour = "black", size = 12, face = "bold")) + 
    scale_y_continuous(expand = c(0,0)) + 
    labs(x = "Pair treatment over time", y = "Relative Abundance (%)", fill = "20 most abundance OTUs") + 
    scale_fill_manual(values = colours)+
  ggtitle("Marilund 2- Paired samples based on seedling type")

   
```

##Lifestyles over time
```{r}
dt <- read.table("C:/Users/salz0002/Documents/R/Lifestyle_relativeabundances.csv", header = TRUE, sep = ",")
```
```{r}
dt$Weeks <- factor(dt$Weeks, levels=dt$Weeks)
```
```{r}
dfm= melt(dt, na.rm = FALSE, value.name = "value")
```

```{r}
colours=c("royalblue","cyan", "lightgreen", "magenta","seagreen","chocolate1", "plum4", "slategray4", "darkgoldenrod2", "darkmagenta", "plum", "snow", "olivedrab", "lightsalmon","lightsalmon4", "lightblue","slateblue", "plum4", "wheat3", "blue", "yellow", "slategray4")
```

```{r}
tax20 = ggplot(dfm, aes(x = Weeks, fill = variable, y = value)) + 
    geom_bar(stat = "identity", colour = "black")+
    theme(axis.text.x = element_text(angle = 90, size = 14, colour = "black", vjust = 0.5, hjust = 1, face= "bold"), 
    axis.title.y = element_text(size = 16, face = "bold"),  axis.title.x = element_text(size = 16, face = "bold"), legend.title = element_text(size = 16, face = "bold"),     panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
    legend.text = element_text(size = 12, face = "bold", colour = "black"), 
    axis.text.y = element_text(colour = "black", size = 12, face = "bold")) + 
    scale_y_continuous(expand = c(0,0)) + 
    labs(x = "Lifest over time", y = "Relative Abundance (%)", fill = "Lifestyles") + 
    scale_fill_manual(values = colours)
```

