---
title: "Master thesis analysis"
author: "Sara"
date: "15/3/2023"
output: html_document
---
```{r}
data <- read.table("C:/Users/saral/OneDrive/Documentos/Bioinformatic/Community_analysis_R/ThesisCommunity_data/CommunityData3.csv", header = TRUE, sep = ";")

library("magrittr")
data<-data %>% na.omit()
```
###Plots over time distribution
##Sites
#Gavtshagen
```{r Split the data or sorf data based on site}
G_site <- data[data$Site=="G",]
G<-subset(G_site, select = c(Site, Treatment, Year, week, reads, OTUs))
G_natural<- G[G$Treatment=="N",]
G_planted<- G[G$Treatment=="P",]

library("ggplot2")
ggplot()+
  geom_line(data = G_planted, mapping = aes(week, reads), color="orange")+
  geom_point(data = G_planted, mapping = aes(week, reads), color="orange")+
  geom_line(data = G_natural, mapping = aes(week, reads), color="darkgreen")+
  geom_point(data = G_natural, mapping = aes(week, reads), color="darkgreen")

#M1
#Split the data or sorf data based on site
M1_site <- data[data$Site=="M1",]
M1<-subset(M1_site, select = c(Site, Treatment, Year, week, reads, OTUs))
M1_natural<- M1[M1$Treatment=="N",]
M1_planted<- M1[M1$Treatment=="P",]

ggplot()+
  geom_line(data = M1_planted, mapping = aes(week, reads), color="orange")+
  geom_point(data = M1_planted, mapping = aes(week, reads), color="orange")+
  geom_line(data = M1_natural, mapping = aes(week, reads), color="darkgreen")+
  geom_point(data = M1_natural, mapping = aes(week, reads), color="darkgreen")

#M2. Split the data or sorf data based on site
M2_site <- data[data$Site=="M2",]
M2<-subset(M2_site, select = c(Site, Treatment, Year, week, reads, OTUs))
M2_natural<- M2[M2$Treatment=="N",]
M2_planted<- M2[M2$Treatment=="P",]

ggplot()+
  geom_line(data = M2_planted, mapping = aes(week, reads), color="orange")+
  geom_point(data = M2_planted, mapping = aes(week, reads), color="orange")+
  geom_line(data = M2_natural, mapping = aes(week, reads), color="darkgreen")+
  geom_point(data = M2_natural, mapping = aes(week, reads), color="darkgreen")
```

###Venn diagram plots
#Sites
```{r}
##Split the data or sorf data based on site
G_site <- data[data$Site=="G",]
g_otu<- subset(G_site, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs))
G_site <- colSums(g_otu)

M1_site <- data[data$Site=="M1",]
M1_otu<- subset(M1_site, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs))
M1_site <- colSums(M1_otu)

M2_site <- data[data$Site=="M2",]
M2_otu<- subset(M2_site, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs))
M2_site <- colSums(M2_otu)

#combine the three subdataframes
library("dplyr")
df <- cbind(G_site,M1_site,M2_site)

df <- as.data.frame(df)

df<-df[rowSums(df) !=0,]

#make binary the dataframe
G<-replace(df$G_site, df$G_site>0,1)

M1<-replace(df$M1_site, df$M1_site>0,1)

M2<-replace(df$M2_site, df$M2_site>0,1)

gsum <-  sum(G) #just to have a look at the number

m1sum <- sum(M1)

m2sum <- sum(M2)
 
df2 <- as.data.frame(cbind(G, M1, M2)) #only for 5 sets

#Draw the diagram
library(limma)

a <- vennCounts(df2)

vennDiagram(a)

##Treatment, natural or planted
#Split the data or sorf data based on site
N_treat <- data[data$Treatment=="N",]
N_otu<- subset(N_treat, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs))
N_treat <- colSums(N_otu)

P_treat <- data[data$Treatment=="P",]
P_otu<- subset(P_treat, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs))
P_treat <- colSums(P_otu)

#combine the three subdataframes
library("dplyr")
df <- cbind(N_treat,P_treat)

df <- as.data.frame(df)

df<-df[rowSums(df) !=0,]

#make binary the dataframe

N<-replace(df$N_treat, df$N_treat>0,1)

P<-replace(df$P_treat, df$P_treat>0,1)

Nsum <-  sum(N) #just to have a look at the number

Psum <- sum(P)

df2 <- as.data.frame(cbind(N, P)) #only for 5 sets

#Draw the diagram
library(limma)

b <- vennCounts(df2)

vennDiagram(b)
```

```{r Year}
##Split the data or sorf data based on site
Y21 <- data[data$Year=="21",]
Y21_otu<- subset(Y21, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs))
Y21 <- colSums(Y21_otu)

Y22<- data[data$Year=="22",]
Y22_otu<- subset(Y22, select=-c(Tag, Site, Sample, Treatment, Year, Info, week, Time_point, reads, OTUs))
Y22 <- colSums(Y22_otu)

#combine the three subdataframes
library("dplyr")
df <- cbind(Y21,Y22)

df <- as.data.frame(df)

df<-df[rowSums(df) !=0,]

#make binary the dataframe

Y21<-replace(df$Y21, df$Y21>0,1)

Y22<-replace(df$Y22, df$Y22>0,1)

y21sum <-  sum(Y21) #just to have a look at the number

y22sum <- sum(Y22)

df2 <- as.data.frame(cbind(Y21, Y22)) #only for 5 sets

#Draw the diagram
library(limma)

c <- vennCounts(df2)

vennDiagram(c)
```

###Normality and homocedasticity test
```{r}
#Histogram
hist(all_data$reads)

hist(all_data$OTUs)
```
##Contraste de hipótesis: normalidad

shapiro.test(all_data$OTUs)
shapiro.test(all_data$reads)

##homocedasticidad
fligner.test(OTUs~Treatment, all_data)

fligner.test(OTUs~Site, all_data)

fligner.test(OTUs~Year, all_data)

fligner.test(OTUs~week, all_data)


fligner.test(reads~Treatment, all_data)

fligner.test(reads~Site, all_data)

fligner.test(reads~Year, all_data)

fligner.test(reads~week, all_data)

#Mis datos son normales y homocedasticos. Heterodascity. The variability is equally distributed--> it doesn't matter the value of the residual or the real value.
```

###Richness test
```{r}
n_otu <- data[11:length(data)]

n_meta <- subset(data[1:10])

str(n_meta)

 library(tidyverse) # great for averaging data and getting standard error for plotting
  library(dplyr)


n_meta$Treatment <- as.factor(n_meta$Treatment)
n_meta$Site <- as.factor(n_meta$Site)
n_meta$week <- as.factor(n_meta$week)
n_meta$Year <- as.factor(n_meta$Year)

all_data <- cbind(n_meta, n_otu)


###Normalised data with read count size.Root square of each OTU per sample

all_data<-cbind(all_data, sqrt(rowSums(n_otu)))
colnames(all_data)[colnames(all_data) =="sqrt(rowSums(n_otu))"]<- "read.counts"

#segun miguelito tienes que chequear las cosas despues

#alfa-diversty --> Shannon
library("vegan")
shann <- diversity(n_otu, "shannon")
simp <- diversity(n_otu, "simpson")


####models
##Treatment
H<-lm(shann~read.counts+Treatment, data = all_data)
H
summary(H)


one<- aov(H, data = all_data)

summary(one)
plot(one)

##Site

all_data$Site <- gsub("G", "0", all_data$Site)
all_data$Site <- gsub("M1", "1", all_data$Site)
all_data$Site <- gsub("M2", "2", all_data$Site)

Hs<-lm(shann~all_data$read.counts + all_data$Site)
Hs
Hs<-lm(shann~all_data$read.counts + as.numeric(all_data$Site))
summary(Hs)

Hs<- aov(Hs, data = all_data)

summary(Hs)

##Year

H<-lm(shann~read.counts+Year, data = all_data)
H

one<- aov(H, data = all_data)

summary(one)

##Week
H<-lm(shann~read.counts+week, data = all_data)
H

one<- aov(H, data = all_data)

summary(one)


####Beta diversity.
#I need to transform the data with Hellinger transformation
library("labdsv")
hell<-hellinger(n_otu)

##Treatment
bd<-adonis2(hell~read.counts+Treatment, data = all_data)
bd

##Site
bd<-adonis2(hell~read.counts+Site, data = all_data)
bd

##Year

bd<-adonis2(hell~read.counts+Year, data = all_data)
bd

##Week
bd<-adonis2(hell~read.counts+week, data = all_data)
bd
```


